Final Report: Scenario 2 - MPX in the EU

### 2022-11-28 MS5.Rmd
Cleaned the methods section and inserted the discussion section.

As the world begins to recover from the effects of COVID, the threat of the monkeypox virus (MPX) has public health officials preparing for the worst. Throughout Europe, cases of MPX have begun to rise and it is imperative to understand the groups at highest risk. This report aims to prepare data to analyze and assess the impact of MPX
within each country of the EU, and understand how case rates may differ by region and other various demographic factors. These findings will aid the implementation and appropriate response to the epidemic to local
health departments. Our team hypothesizes that monkeypox will be prevalent among the very young and old as is common for many infectious diseases.

@@

Datasets with information on confirmed cases of monkeypox per country as well as census data from Europe were used to prepare this report. Countries were grouped by region based on world_country_region data set to describe the cases per region.

Within the euro_mpx_cases dataset, relevant fields include country, country code, and confirmed cases. A curated demographic dataset included information on age, sex, and number of people per strata.

@@

In table 1, we see that the regions in EU include various numbers of countries, thus
having unequal populations. A useful metric to create a graph will be
the calculation of cases per 100,000. This will allow us to use 1
categorical variable (EU Region) to a quantitative variable (number of
confirmed cases). We compiled the relevant confirmed case information per region in the
EU.

@@

In the bar graph (fig. 1) above we see the breakdown of monthly MPX infections in Europe between May and August of this year. The majority of cases occurred in Western Europe and Southern Europe as evidenced by the respectively colored purple and green regions in the bar graph. Over the course of 4 months, the total number of MPX cases by region were Western Asia = 4, Eastern Europe = 277, Northern Europe = 619, Southern Europe = 7958, and Western Europe = 8214

@@

As a follow up to the bar graph (fig. 1) above, we wanted to look at the rate of MPX infections per region. In the bar graph above, Western Europe had the greatest number of MPX infections however when we look at the rate of infections (fig. 2), we see that over the course of the 4 months, Southern Europe had the greatest rate of infection with a peak of 2.84 MPX cases per 100,000 people in the month of July.

@@

>>>>

As the world begins to recover from the effects of COVID, the threat of the so-called monkeypox virus (MPX) has public health officials preparing for the worst. Throughout Europe (EU), cases of MPX have begun to rise, and it is imperative to understand the groups at the highest risk for infection. This report aims to prepare data to analyze and assess the impact of MPX within each country of the EU, understand how case rates may differ by region, and find potential links to various demographic factors. These findings will aid local health departments' implementation and appropriate response to the epidemic. Our team hypothesizes that MPX will be prevalent among young (under 30 years old) and older populations (over 64 years old), as is typical for many infectious diseases.

@@

To prepare this report, use data sets for the EU, explicitly including information on confirmed cases of MPX per country (euro_mpx_cases.csv) and census data (euro_census_stats.csv). The trajectory of this investigation incorporated four sets of data manipulation:
(1) Grouping countries into regions;
(2) Assessing necessary census information;
(3) Compiling data sets with confirmed case demographics;
(4) Creating tables and figures for visual analysis.
Countries were grouped by region based on a region's data set (world_country_regions.csv) to enhance the representation of positive cases per region. Within the euro_mpx_cases data set, relevant fields include country, country code, and confirmed cases. A curated demographic data set included information on age, sex, and the number of people per strata. 
As an aside with the census data, we chose to initially assess any "missing" data as "NA" when paired with a country. However then decided to print information with omit-NA's to allow aggregate fields to be added together for "Table 2: Population Demographic (Population Density) per Region."

@@

In Table 1, we see that the regions in EU include various numbers of countries, thus
having unequal populations. A useful metric to create a graph will be
the calculation of cases per 100,000. This will allow us to use one
categorical variable (EU Region) to a quantitative variable (number of
confirmed cases). We compiled the relevant confirmed case information per region in the
EU.

@@

In the bar graph (Fig. 1) above we see the breakdown of monthly MPX infections in Europe between May and August of this year. The majority of cases occurred in Western Europe and Southern Europe as evidenced by the respectively colored purple and green regions in the bar graph. Over the course of 4 months, the total number of MPX cases by region were Western Asia = 4, Eastern Europe = 277, Northern Europe = 619, Southern Europe = 7958, and Western Europe = 8214

@@

As a follow up to the bar graph (Fig. 1) above, we wanted to look at the rate of MPX infections per region. In the bar graph above, Western Europe had the greatest number of MPX infections however when we look at the rate of infections (Fig. 2), we see that over the course of the four months, Southern Europe had the greatest rate of infection with a peak of 2.84 MPX cases per 100,000 people in the month of July.

@@

Our analysis of Table 1 and Figure 1 reveal the impacted regions for MPX though the study period (May through August 2022) occurred in the Southern and Western Regions of Europe. Within the EU, these two regions comprise 13 of its 29 total countries, roughly 71.8% of the total EU population (extrapolated using Table 1).

Through empirical analysis in Figure 3, we are not confident of the initial hypothesis regarding younger (under 30 years old) and older (over 64 years old) populations having a higher risk of infection. We observe that the age group for 30-64-year-olds (broken into two categories of 30-49 and 50-64 years) contributes over 50% of MPX cases in countries with a minimum of 100 confirmed MPX cases (six countries from the 29 countries in the EU). Given that the 30-64 age group appears to burden more patients than those below 29 years old and above 64 years old for all six of the countries that have 100+ MPX cases, we believe that the expansion of this finding will be evident throughout the remaining 23 countries in the EU.

Given the analysis of this information, health department leadership should begin educational programs and outreach focusing on those 30-64 aged individuals.
